---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

# If emo is not installed, install it.
# devtools::install_github("hadley/emo")
```

# meetupr
[![Build Status](https://travis-ci.org/rladies/meetupr.svg?branch=master)](https://travis-ci.org/rladies/meetupr)

R interface to the [Meetup API](https://www.meetup.com/meetup_api/) (v3)

## Installation

To install the development version from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rladies/meetupr")
```
A released version will be on CRAN [soon](https://github.com/rladies/meetupr/issues/24).

## Usage

### Authentication

In order to use this package, you can use our built-in [OAuth](https://www.meetup.com/meetup_api/auth/) credentials (recommended), or if you prefer, you can supply your own by setting the `meetupr.consumer_key` and `meetupr.consumer_secret` variables. 

Each time you use the package, you will be prompted to log in to your meetup.com account.  The first time you run any of the **meetupr** functions in your session, R will open a browser window, prompting you to "Log In and Grant Access" (to the **meetupr** "application").

_Note: As of August 15, 2019, Meetup.com switched from an API key based authentication system to OAuth 2.0, so we [added support](https://github.com/rladies/meetupr/issues/51) for OAuth.  For backwards compatibility, the functions all still have a `api_key` argument which is no longer used and will eventually be [deprecated](https://github.com/rladies/meetupr/issues/59)._ 

### Functions

We currently have the following functions:

* `get_members()`
* `get_boards()`
* `get_events()`  
* `get_event_attendees()`  
* `get_event_comments()`
* `get_event_rsvps()`
* `find_groups()`

Each will output a tibble with information extracted into from the API as well as a `list-col` named `*_resource` with all API output. For example, the following code will get all upcoming events for the [R-Ladies San Francisco](https://meetup.com/rladies-san-francisco) meetup.

```{r}
library(meetupr)

urlname <- "rladies-san-francisco"
events <- get_events(urlname, "past")
dplyr::arrange(events, desc(created))
```

Next we can look up all R-Ladies groups by "topic id".  You can find topic ids for associated tags by querying [here](https://secure.meetup.com/meetup_api/console/?path=/find/topics).  The `topic_id` for topic, "R-Ladies", is `1513883`.

```{r}
groups <- find_groups(topic_id = 1513883)
dplyr::arrange(groups, desc(created))
```


## How can you contribute?

We are looking for new people to join the list of contributors!  Please take a look at the open [issues](https://github.com/rladies/meetupr/issues), file a new issue, contribute tests, or improve the documentation.  We are also looking to expand the set of functions to include more endpoints from the [Meetup API](https://www.meetup.com/meetup_api/).  Lastly, we'd also love to [hear about](https://github.com/rladies/meetupr/issues/74) any applications of the **meetupr** package, so we can compile a list of demos!

Please note that the this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
