---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

# devtools::install_github("hadley/emo")
```

# meetupr
[![Build Status](https://travis-ci.org/rladies/meetupr.svg?branch=master)](https://travis-ci.org/rladies/meetupr)

R interface to the Meetup API (v3)

**Authors:** [Gabriela de Queiroz](http://gdequeiroz.github.io/), [Erin LeDell](http://www.stat.berkeley.edu/~ledell/), [Olga Mierzwa-Sulima](https://github.com/olgamie), [Lucy D'Agostino McGowan](http://www.lucymcgowan.com), [Claudia Vitolo](https://github.com/cvitolo)<br/> [MIT](https://opensource.org/licenses/MIT)
**License:** [MIT](https://opensource.org/licenses/MIT)

## Installation

To install the development version from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("rladies/meetupr")
```
A released version will be on CRAN soon.

## Usage

To use this package, you will first need to get your meetup API key. To do so, go to this link: [https://secure.meetup.com/meetup_api/key/](https://secure.meetup.com/meetup_api/key/)

Once you have your key, save it as an environment variable by running the following:

```{r, eval = FALSE}
Sys.setenv(MEETUP_KEY = "PASTE YOUR MEETUP KEY HERE")
```

If you don't want to save it here, you can input it in each function using the `api_key` parameter (just be sure not to send any documents with your key to GitHub `r emo::ji("speak_no_evil")`).

We currently have the following functions:

* `get_members()`
* `get_boards()`
* `get_events()`  
* `get_event_attendees()` *Note: Must be a Meetup group administrator to use because of API restriction.
* `get_event_comments()`
* `get_event_rsvps()` 
* `find_groups()`

Each will output a tibble with information extracted into from the API as well as a `list-col` named `*_resource` with all API output. For example, the following code will get all past events for the [RLadies Nashville](https://meetup.com/rladies-nashville) meetup.

```{r}
library(meetupr)

urlname <- "rladies-nashville"
(events <- get_events(urlname, event_status = "past"))
```


## How can you contribute?

Take a look at some resources:

- [https://www.meetup.com/meetup_api/](https://www.meetup.com/meetup_api/)
- [https://www.meetup.com/meetup_api/clients/](https://www.meetup.com/meetup_api/clients/)

In order to run our tests, you will have to set the `urlname` for meetup you belong to as an environment variable using the following code:
```{r, eval = FALSE}
Sys.setenv(MEETUP_NAME = "YOUR MEETUP NAME")
```

### TO DO:
- add tests

