---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

# If emo is not installed, install it.
# devtools::install_github("hadley/emo")
```

# meetupr
[![Build Status](https://travis-ci.org/rladies/meetupr.svg?branch=master)](https://travis-ci.org/rladies/meetupr)

R interface to the Meetup API (v3)

**Authors:** [Gabriela de Queiroz](http://gdequeiroz.github.io/), [Erin LeDell](http://github.com/ledell/), [Olga Mierzwa-Sulima](https://github.com/olgamie), [Lucy D'Agostino McGowan](http://www.lucymcgowan.com), [Claudia Vitolo](https://github.com/cvitolo)<br/> 
**License:** [MIT](https://opensource.org/licenses/MIT)

## Installation

To install the development version from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("rladies/meetupr")
```
A released version will be on CRAN [soon](https://github.com/rladies/meetupr/issues/24).

## Usage

### Authentication

As of August 15, 2019, Meetup.com switched from an API key based authentication system to [OAuth 2.0](https://www.meetup.com/meetup_api/auth/), so we now [support OAuth authentication](https://github.com/rladies/meetupr/issues/51).  The functions all have a `api_key` argument which is no longer used and will eventually be [deprecated](https://github.com/rladies/meetupr/issues/59).  In order to use this package, you can use our built-in OAuth credentials (recommended), or if you prefer, you can supply your own by setting the `meetupr.consumer_key` and `meetupr.consumer_secret` variables. 

Each time you use the package, you will be prompted to log in to your meetup.com account.  The first time you run any of the **meetupr** functions in your session, R will open a browser window, prompting you to "Log In and Grant Access" (to the **meetupr** "application").

### Functions

We currently have the following functions:

* `get_members()`
* `get_boards()`
* `get_events()`  
* `get_event_attendees()`  
* `get_event_comments()`
* `get_event_rsvps()`
* `find_groups()`

Each will output a tibble with information extracted into from the API as well as a `list-col` named `*_resource` with all API output. For example, the following code will get all upcoming events for the [R-Ladies San Francisco](https://meetup.com/rladies-san-francisco) meetup.

```{r}
library(meetupr)

urlname <- "rladies-san-francisco"
events <- get_events(urlname, "past")
dplyr::arrange(events, desc(created))
```

Next we can look up all R-Ladies groups by "topic id".  You can find topic ids for associated tags by querying [here](https://secure.meetup.com/meetup_api/console/?path=/find/topics).  The `topic_id` for topic, "R-Ladies", is `1513883`.

```{r}
groups <- find_groups(topic_id = 1513883)
dplyr::arrange(groups, desc(created))
```


## How can you contribute?

Take a look at some resources:

- [https://www.meetup.com/meetup_api/](https://www.meetup.com/meetup_api/)
- [https://www.meetup.com/meetup_api/clients/](https://www.meetup.com/meetup_api/clients/)

We are looking for people to help write tests and vignettes!  You can also take a look at the open [issues](https://github.com/rladies/meetupr/issues).

In order to run our tests, you will have to set the `urlname` for meetup you belong to as an environment variable using the following code:
```{r, eval = FALSE}
Sys.setenv(MEETUP_NAME = "YOUR MEETUP NAME")
```


--

Please note that the 'meetupr' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
