---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the Rmd file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

# If emo is not installed, install it.
# devtools::install_github("hadley/emo")
```

# meetupr
<!-- badges: start -->
[![R-CMD-check](https://github.com/rladies/meetupr/workflows/R-CMD-check/badge.svg)](https://github.com/rladies/meetupr/actions)
[![Codecov test coverage](https://codecov.io/gh/rladies/meetupr/branch/master/graph/badge.svg)](https://codecov.io/gh/rladies/meetupr?branch=master)
<!-- badges: end -->

R interface to the [Meetup API](https://www.meetup.com/meetup_api/) (v3)

## Installation

To install the development version from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rladies/meetupr")
```
A released version will be on CRAN [soon](https://github.com/rladies/meetupr/issues/24).

## Usage

### Authentication

#### API key? No

```{r child='man/rmd-fragments/api-key.Rmd'} 
```

#### OAuth? Yes

```{r child='man/rmd-fragments/oauth.Rmd'} 
```

### Functions

See the [pkgdown reference](https://rladies.github.io/meetupr/reference/index.html).

For example, the following code will get all upcoming events for the [R-Ladies San Francisco](https://meetup.com/rladies-san-francisco) meetup.

```{r}
library(meetupr)

urlname <- "rladies-san-francisco"
events <- get_events(urlname, "past")
dplyr::arrange(events, desc(created))
```

Next we can look up all R-Ladies groups by "topic id".  You can find topic ids for associated tags by querying [here](https://secure.meetup.com/meetup_api/console/?path=/find/topics).  The `topic_id` for topic, "R-Ladies", is `1513883`.

```{r}
groups <- find_groups(topic_id = 1513883)
dplyr::arrange(groups, desc(created))
```


## How can you contribute?

We are looking for new people to join the list of contributors!  Please take a look at the open [issues](https://github.com/rladies/meetupr/issues), file a new issue, contribute tests, or improve the documentation.  We are also looking to expand the set of functions to include more endpoints from the [Meetup API](https://www.meetup.com/meetup_api/).  Lastly, we'd also love to [hear about](https://github.com/rladies/meetupr/issues/74) any applications of the **meetupr** package, so we can compile a list of demos!

Please note that the this project is released with a [Contributor Code of Conduct](https://github.com/rladies/.github/blob/master/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
