---
title: "R-Ladies chapters on meetup.com"
author: "Claudia Vitolo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  eval = FALSE
)
```

## Install and load packages

```{r install, eval = FALSE}
devtools::install_github("rladies/meetupr")

# Set up your meetup API key
Sys.setenv(MEETUP_KEY = "PASTE YOUR MEETUP KEY HERE")
```

```{r load}
library("meetupr")
library("purrr")
library("dplyr")
library("lubridate")
library("tidyr")
library("forecast")
library("ggplot2")
```

## How many R-Ladies chapter are out there?

```{r groups}
meetup_groups <- find_groups(text = "r-ladies")

# Keep only groups whose name starts with "r"
first_letter <- tolower(substr(meetup_groups$name, 1, 1))
meetup_groups <- meetup_groups[which(first_letter == "r"),]

# Keep only groups whose urlname contains the string "ladies"
urlnames <- meetup_groups$urlname
meetup_groups <- meetup_groups[which(grepl("ladies", tolower(urlnames))),]

rm(first_letter, urlnames)
```

## Is the urlname consistent across groups?

The official urlname should follow this naming convention:
__rladies-cityname__.

*How many chapters have a consistent urlname?*

```{r groupA}
urlnames <- meetup_groups$urlname

groupA <- urlnames[grep(pattern = "rladies-",
                              x = urlnames,
                              ignore.case = FALSE)]
groupA
```

*How many chapters have a problem with capitalisation?*

```{r groupB}
groupB <- urlnames[grep(pattern = "rladies-",
                              x = urlnames,
                              ignore.case = TRUE)]
groupB <- setdiff(groupB, groupA)
groupB
```

Here is the list of chapters that need some extra changes:

```{r groupC}
groupC <- setdiff(urlnames, c(groupA, groupB))
groupC
```

## Is the chapter name consistent across groups?

The official chapter name should follow this naming convention:
__R-Ladies Cityname__.

*How many chapters have a consistent name?*

```{r groupD}
chapternames <- meetup_groups$name
groupD <- chapternames[grep(pattern = "R-Ladies ",
                            x = chapternames,
                            ignore.case = FALSE)]
groupD
```

The chapters below should, ideally, change their names

```{r groupE}
groupE <- setdiff(chapternames, groupD)
groupE
```

## Chapters to migrate to meetup pro

*Which chapters do not belong to R-Ladies Global?*

```{r migration}
chapters2migrate <- meetup_groups[meetup_groups$organizer != "R-Ladies Global",]
chapters2migrate$name
```

## Chapter activity

To check chapter activity is better to use R-Ladies Global API key so that we have access to private groups.

Let's now extract the number of past events and the dates of last and next events, for each chapter.

```{r activity}
meetup_groups$past_events <- NA
meetup_groups$date_last_event <- NA
meetup_groups$date_next_event <- NA

# To avoid to exceed API request limit we use Jesse Maegan's solution:
# https://github.com/rladies/meetupr/issues/30
slowly <- function(f, delay = 0.5) {
  function(...) {
    Sys.sleep(delay)
    f(...)
  }
}
x <- map(meetup_groups$urlname,
         slowly(safely(get_events)),
         event_status = c("past", "upcoming"))

for (i in seq_along(x)){
  message(meetup_groups$urlname[i])
  if (is.null(x[[i]]$error)){
    temp <- x[[i]]$result
    if ("past" %in% temp$status){
      past <- temp %>%
        group_by(status) %>%
        summarise(count = n(),
                  mydates = format(max(local_date), '%Y-%m-%d'))
    meetup_groups$past_events[i] <- past$count[past$status == "past"]
    meetup_groups$date_last_event[i] <- past$mydates[past$status == "past"]
    }else{
      meetup_groups$past_events[i] <- 0
      meetup_groups$date_last_event[i] <- NA
    }
    if ("upcoming" %in% temp$status){
      upcoming <- temp %>%
        group_by(status) %>%
        summarise(count = n(),
                  mydates = format(min(local_date), '%Y-%m-%d'))
      meetup_groups$date_next_event[i] <- upcoming$mydates[upcoming$status == "upcoming"]
    }else{
      meetup_groups$date_next_event[i] <- NA
    }
  }else{
    meetup_groups$past_events[i] <- 0
    meetup_groups$date_last_event[i] <- NA
    meetup_groups$date_next_event[i] <- NA
  }
}
```

*Which chapters have never had events and have not even planned one (and have been created more than 6 months ago)?*

```{r noevents}
no_events <- meetup_groups %>%
  filter(past_events == 0,
         is.na(date_next_event),
         created < as.POSIXct("2018-04-30")) %>%
  arrange(created)
```

*Which chapters had no events in the past 6 months and have not even planned one?*

```{r no_recent_events}
no_recent_events <- meetup_groups %>%
  filter(date_last_event < as.POSIXct("2018-04-30"),
         is.na(date_next_event)) %>%
  arrange(date_last_event)
```

## R-Ladies quarterly growth
Estimated quarterly growth of chapter:

```{r ygrowth}
# Expand dataframe by quarter
quarterly_chapters <- meetup_groups %>%
  mutate(Date = as.Date(created, "%Y-%m-%d")) %>%
  complete(Date = seq.Date(min(Date), max(Date), by="day")) %>%
  group_by(qrtr = floor_date(Date, "quarter")) %>%
  summarize(count = sum(!is.na(urlname))) %>%
  mutate(cumsum = cumsum(count))

# Forecast next year's active chapters
chap <- ts(data = quarterly_chapters$cumsum, start = c(2012, 4), frequency = 4)
fc <- holt(y = chap, h = 5, damped = TRUE)

autoplot(chap) +
  autolayer(fc, series="Holt's method", PI=TRUE) +
  ggtitle("2019 forecast using Holt's method") + xlab("") +
  ylab("R-Ladies active chapters") +
  guides(fill=FALSE, color=FALSE) + theme_bw()

average_chap_nexty <- round(fc$upper[3], 0)

round(fc$upper[c(3, 5, 7, 9)], 0)
```

## R-Ladies yearly budget

The quarterly budget is calculated as the sum of fixed costs and variable costs, where the variable costs are proportional to the expected number of chapters in the next quarter.

```{r ybudget}
# Variable costs (proportional to the number of chapters: meetup-pro fees, merchandise and shipping costs
variable_costs <- 230 * average_chap_nexty

# Fixed costs:
# Web/email hosting, domain registration, mattermost, nonprofit filing
fixed_costs <- 1188

# Yearly budget
fixed_costs + variable_costs
```
