---
title: "Setting up and Getting Started with meetupr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting up and Getting Started with meetupr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
#| label: setup

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)

vcr::setup_knitr(prefix = "meetupr-")
#meetupr:::mock_if_no_auth()
Sys.setenv(MEETUPR_DEBUG = "")
```

```{r}
#| label: load
library(meetupr)
```

## Quick Setup Check

Start by running `meetup_sitrep()` to set up authentication or check your authentication status:

```{r}
#| label: sitrep
#| cassette: true

meetup_sitrep()
```

This function will show you:  
- Which authentication method is active  
- Whether your credentials are properly configured  
- API connectivity status  
- Setup instructions if authentication is missing  

## Authentication Methods

meetupr supports two authentication methods:

### OAuth Authentication (Interactive)

Best for interactive development, data exploration, and one-off analyses.

**Setup:**
```r
# Authenticate interactively (opens browser)
meetup_auth()

# Check authentication status
meetup_auth_status()
```

The OAuth token is cached locally in your httr2 cache directory. 
Token files follow the pattern: `{hash}-token.rds.enc`

### CI/CD Authentication (Non-Interactive)

For automated workflows, you can set up token-based authentication:

**1. Authenticate locally first:**
```r
meetup_auth()
meetup_auth_setup_ci()
```

This encodes your token as base64 and provides instructions for setting CI secrets.

**2. Set environment variables in your CI:**
- `MEETUP_TOKEN` - Base64-encoded token (from step 1)
- `MEETUP_TOKEN_FILE` - Token filename (e.g., `ae743e0-token.rds.enc`)

**3. Token loads automatically in CI** 

To make the authentication work in CI/CD pipelines, add this line before running any API calls:

```r
meetup_auth_load_ci()
```



## Environment Variables Reference

| Variable | Required | Purpose |
|----------|----------|---------|
| `MEETUP_CLIENT_ID` | Optional* | OAuth client ID |
| `MEETUP_CLIENT_SECRET` | Optional* | OAuth client secret |
| `MEETUP_CLIENT_NAME` | No | OAuth client name (default: "meetupr") |
| `MEETUP_TOKEN` | for CI | Base64-encoded OAuth token |
| `MEETUP_TOKEN_FILE` | for CI | Token filename |
| `MEETUPR_DEBUG` | for CI | Enable debug output ("1" = on) |

\* Built-in credentials are used if not provided, and Meetup clients are only available for approved apps from Pro (paid) accounts.

### Setting Environment Variables

If you want to use your own Meetup OAuth app credentials, set them as environment variables.
Use `.Renviron` for local development:

```r
usethis::edit_r_environ()
```

Add:

```bash
MEETUP_CLIENT_ID=your_client_id
MEETUP_CLIENT_SECRET=your_secret
```

The `.Renviron` file contains potentially sensitive credentials.
While the client ID and secret are not as sensitive as the token, treat them with care.
However, despite secrets generally should remain secret, noone can actually use the Meetup API without authenticating as a user, so the risk is lower. 

### Verification

After setting environment variables, restart R and check:

```{r}
#| label: verify-sitrep
#| cassette: true
meetup_sitrep()
```

## Basic Usage Examples

### Get Group Information

```{r}
#| label: group-info
#| cassette: true
group_info <- get_group("rladies-lagos")
str(group_info)
```

### List Upcoming Events

```{r}
#| label: upcoming-events
#| cassette: true
events <- get_events("rladies-lagos")
head(events)
```

### Search for Groups

```{r}
#| label: search-groups
#| cassette: true
r_groups <- find_groups("R programming")
head(r_groups)
```

## Debug Mode

Sometimes issues arise due to misconfiguration.
Enabling debugging mode provides more information on what is sent to the API, i.e. the full graphql query.
Enable debugging mode by setting system variable "MEETUPR_DEBUG" to 1 (0 or missing means no debugging):

```{r}
#| label: debug
#| cassette: true

local_meetupr_debug(1)
meetup_sitrep()
find_groups("R programming")
```

to turn it off again, set it to 0 or unset it:

```{r}
#| label: debug-off
#| cassette: true

local_meetupr_debug(0)
meetup_sitrep()
find_groups("R programming")
```

## Common Issues

### Authentication Errors

```r
Error: Authentication required
```

**Solution**: Run `meetup_sitrep()` to diagnose authentication issues.

### Invalid RSA Key

```r
Error: RSA private key not found
```

**Solution**: Verify your RSA key format includes header/footer lines and is properly formatted.

### Rate Limiting

```r
Error: Rate limit exceeded
```

**Solution**: Implement delays between requests or use caching strategies.

## Next Steps

- **For R-Ladies chapter management**: See [R-Ladies vignette](rladies.html)  
- **For custom GraphQL queries**: See [GraphQL vignette](graphql.html)`



