---
title: "Find and analyse R-Ladies chapters on meetup.com"
author: "Claudia Vitolo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  eval = FALSE
)
```

## Install and initialise package

```{r install, eval = FALSE}
devtools::install_github("rladies/meetupr")
```

```{r load}
library("meetupr")

api_key <- "YOUR_API_KEY_GOES_HERE"
```

## How many R-Ladies chapter are out there?

```{r groups}
meetup_groups <- find_groups(api_key = api_key, text = "r-ladies")

rladies_groups <- as.character(unlist(meetup_groups$urlname))
# Cleanup
rladies_groups <- rladies_groups[grep(pattern = "ladies", 
                                      x = rladies_groups,
                                      ignore.case = TRUE)]
rladies_groups <- rladies_groups[grep(pattern = "r", 
                                      x = rladies_groups,
                                      ignore.case = TRUE)]
rladies_groups <- rladies_groups[-grep(pattern = "gentlemen",
                                       x = rladies_groups,
                                       ignore.case = TRUE)]
```

Wooowww, there seem to be 59 R-Ladies groups on meetup.com!

## Is the urlname consistent across groups?

The official urlname should follow this naming convention:
__rladies-cityname__. However, only the following chapters have a consistent urlname:

```{r consistency}
excellent <- rladies_groups[grep(pattern = "rladies-", 
                                 x = rladies_groups,
                                 ignore.case = FALSE)]
excellent

# [1] "rladies-barcelona"       "rladies-madrid"          "rladies-london"    
# [4] "rladies-berlin"          "rladies-natal"           "rladies-la"        
# [7] "rladies-munich"          "rladies-montevideo"      "rladies-santa-rosa"
# [10] "rladies-tbilisi"         "rladies-buenos-aires"    "rladies-paris"    
# [13] "rladies-charlottesville" "rladies-nashville"       "rladies-scl"
# [16] "rladies-cape-town"       "rladies-san-diego"       "rladies-montreal" 
# [19] "rladies-philly"          "rladies-dc"              "rladies-portland-maine" 
# [22] "rladies-ldnont"          "rladies-newyork"         "rladies-chicago"
# [25] "rladies-istanbul"        "rladies-san-francisco"   "rladies-connecticut"
# [28] "rladies-austin"          "rladies-san-jose"        "rladies-adelaide"
```

Few only have a problem with capitalisation:
```{r good}
good <- rladies_groups[grep(pattern = "rladies-",
                            x = rladies_groups,
                            ignore.case = TRUE)]
good <- setdiff(good, excellent)
good

# [1] "RLadies-Columbus" "RLadies-Buffalo"
```

Here is the list of chapter that got it almost right:
```{r almost}
almost <- setdiff(rladies_groups, c(excellent, good))
almost

# [1] "R-Ladies-Belgrade"       "R-Ladies_Quito"          "R-Ladies-Boston" 
# [4] "R-Ladies-Dublin"         "R-Ladies-Ames"           "R-Ladies-Manchester"
# [7] "R-Ladies-Vancouver"      "R-Ladies-Melbourne"      "R-Ladies-Brussels"
# [10] "R-Ladies-Miami"          "R-Ladies-Orlando"        "R-Ladies-Strasbourg"
# [13] "R-Ladies-RTP"            "R-Ladies-Sarasota"       "R-Ladies-Milan"
# [16] "R-Ladies-Taipei"         "R-Ladies-Lisboa"         "R-Ladies-St-Louis"  
# [19] "R-Ladies-Americana"      "R-Ladies-PDX"            "R-Ladies-Budapest"
# [22] "R-Ladies-Chennai"        "R-Ladies-Rio-de-Janeiro" "RLadiesTC"
# [25] "R-Ladies-İzmiR"          "R-Ladies-Bogota"         "rladiescdmx"  
```
