---
title: "Setup for using meetupr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(meetupr)
```

## API key? No

```{r child='../man/rmdhunks/api-key.Rmd'} 
```

## OAuth? Yes

```{r child='../man/rmdhunks/oauth.Rmd'} 
```

## What about running my code on a server

Please read [httr vignette about secrets](https://httr.r-lib.org/articles/secrets.html) and [gargle vignette about managing tokens securely](https://gargle.r-lib.org/articles/articles/managing-tokens-securely.html).

Here is how we handle our secret token for testing the package, it might inspire you.

### Create and encrypt the token

First we create a token. 
We build and git ignore it, even safer would be to put it elsewhere on the computer, 
maybe a temporary file!

```r
token_path <- testthat::test_path(".meetup_token.rds")
use_build_ignore(token_path)
use_git_ignore(token_path)

meetupr::meetup_auth(
  token = NULL,
  cache = TRUE,
  set_renv = FALSE,
  token_path = token_path
)
```

We create a key and save it as an environment variable.
We used an internal function for saving it, you could copy-paste it to .Renviron by hand (see also `usethis::edit_r_environ()`).

```r
# sodium_key <- sodium::keygen()
# set_renv("MEETUPR_PWD" = sodium::bin2hex(sodium_key))
```

Using the key we encrypt the file.
secret.rds will be committed to our repo!

```r
key <- cyphr::key_sodium(sodium::hex2bin(Sys.getenv("MEETUPR_PWD")))

cyphr::encrypt_file(
  token_path,
  key = key,
  dest = testthat::test_path("secret.rds")
)
```

### Use the token on CI

* Save the `MEETUPR_PWD` environment variable as secret on CI (actually called repo secret for e.g. GitHub Actions)

* Use the code below to decrypt the secret before using meetupr.
We save the temporary token to disk but to a temporary folder so it won't end up in a package check artefact.

```r
key <- cyphr::key_sodium(sodium::hex2bin(Sys.getenv("MEETUPR_PWD")))

temptoken <- tempfile(fileext = ".rds")

cyphr::decrypt_file(
  testthat::test_path("secret.rds"),
  key = key,
  dest = temptoken
)

token <- readRDS(temptoken)[[1]]

meetupr::meetup_auth(
  token = token,
  set_renv = FALSE,
  cache = FALSE
) -> token

Sys.setenv(MEETUPR_PAT = temptoken)

```
